---
title: "Exercise 2"
date: "`r Sys.Date()`"
output: 
  bookdown::html_document2
runtime: shiny
---

# Exercise 2

```{r}
library(tidyverse)
library(fitzRoy)
library(cassowaryr)
library(plotly)
library(shiny)
```

```{r}
load("data/aflw_num.rda")
load("data/aflw.rda")
load("data/aflw_scags.rda")
```

## Question a

There are `r n_distinct(aflw$team.name)` teams in the competition. There are `r n_distinct(aflw$player.playerId)` players. There are `r n_distinct(aflw$round.roundNumber)` rounds.

## Question b

```{r team-goals, fig.cap="Goals by team. Frematle and Kangaroos are very close in the number of goals, but Fremantle just edges it"}
aflw %>% 
  group_by(team.name) %>% 
  summarise(goals = sum(goals)) %>% 
  ggplot(aes(goals, reorder(team.name, goals))) +
  geom_col(fill = "#20B2AA") +
  labs(x = "Goals",
       y = "Team",
       title = "Goals by Team") +
  theme_minimal()
```

According to \@ref(fig:team-goals), Fremantle and Kangaroos are neck-to-neck with just over 50 goals scored by each team, but Fremantle is just ahead. Based on the number of goals, **Fremantle** seems to be the team that is most likely to win if the season were completed.

## Question c

A pairs plot of all the numeric variables would create `r scales::comma(34^2)` total plots, of which `r sum(seq(33, 1, -1))` are scatter plots.

## Question d

For each player, the mean of each statistic will be calculated.

```{r}
player_summaries <- aflw %>% 
  group_by(player.playerId) %>% 
  summarise(across(is.numeric, mean)) %>% 
  select(-c(round.roundNumber:player.player.player.playerJumperNumber, player.playerJumperNumber))
```

To find the pair of variables with the "L" shape, the variables pair with the highest **striated2** scagnostic is identified and plotted.

```{r}
aflw_scags %>% 
  arrange(-striated2) %>% 
  head(1)
```

The pair is **hitouts** and **bounces**. We can see the "L" shape below

```{r }
ggplot(player_summaries, aes(hitouts, bounces)) +
  geom_point() +
  theme_classic() +
  labs(title = "Bounces vs Hitouts")
```

The figure shows that players who achieve hitouts almost exclusively never achieve bounces; conversely, players who achieve bounces almost never achieve hitouts.


## question e

To find a pair of variables with a barrier, the variables pair with the highest **monotonic** scagnostic is identified and plotted.

```{r}
aflw_scags %>% 
  arrange(-monotonic) %>% 
  head(1)
```

The pair is **clearances.totalClearances** and **clearances.stoppageClearances**.

```{r }
ggplot(player_summaries, aes(clearances.totalClearances, clearances.stoppageClearances)) +
  geom_point() +
  theme_classic() +
  labs(x = "total clearances",
       y = "stoppage clearances",
       title = "Stoppage Clearances vs Total Clearances")
```

The figure shows that there is barrier of having higher stoppage clearances than total clearances, since it is impossible for that to happen.


## question f


```{r, eval=FALSE}
shinyApp(
ui <- fluidPage(
  plotlyOutput("parcoords"),
  verbatimTextOutput("data")),


server <- function(input, output, session) {
  aflw_numeric <- aflw_scags[,3:15]
  output$parcoords <- renderPlotly({
    dims <- Map(function(x, y) {
      list(values = x, range = c(0,1), label = y)
    }, aflw_numeric, names(aflw_numeric), USE.NAMES = FALSE)
    plot_ly(type = 'parcoords', dimensions = dims,
            source = "pcoords") %>%
      layout(margin = list(r = 30)) %>%
      event_register("plotly_restyle")
  })

  ranges <- reactiveValues()
  observeEvent(event_data("plotly_restyle", source = "pcoords"),
               {
                 d <- event_data("plotly_restyle", source = "pcoords")
                 # what is the relevant dimension (i.e. variable)?
                 dimension <- as.numeric(stringr::str_extract(names(d[[1]]),
                                                              "[0-9]+"))
                 # If the restyle isn't related to a dimension, exit early.
                 if (!length(dimension)) return()
                 # careful of the indexing in JS (0) versus R (1)!
                 dimension_name <- names(aflw_numeric)[[dimension + 1]]
                 # a given dimension can have multiple selected ranges
                 # these will come in as 3D arrays, but a list of vectors
                 # is nicer to work with
                 info <- d[[1]][[1]]
                 ranges[[dimension_name]] <- if (length(dim(info)) == 3) {
                   lapply(seq_len(dim(info)[2]), function(i) info[,i,])
                 } else {
                   list(as.numeric(info))
                 }
               })
  ## filter to the rows that match the selection ranges
  aflw_selected <- reactive({
    keep <- TRUE
    for (i in names(ranges)) {
      range_ <- ranges[[i]]
      keep_var <- FALSE
      for (j in seq_along(range_)) {
        rng <- range_[[j]]
        keep_var <- keep_var | dplyr::between(aflw_scags[[i]],
                                              min(rng), max(rng))
      }
      keep <- keep & keep_var
    }
    aflw_scags[keep, ]
  })
  output$data <- renderPrint({
    tibble::as_tibble(aflw_selected())
  })
},

options = list(height = 600)
)
```

### i.

From the parallel coordinate plot, we can see **positive** associations between some variables such as between *stringy* and *striated*, *striated* and *striated2*, *monotonic* and *splines*, and *splines* and *dcor*. We can observe **negative** associations such as between *clumpy* and *sparse*, *monotonic* and *striated2*, *splines* and *striated2*, and *dcor* and *striated2*. **Outliers** also exist in some variable combinations such as between *clumpy* and *clumpy2*. **Clusters** can also be found in the combination between *striated* and *striated2*.

### ii.

Pairs that have high values on convex tend to have low values on **outlying**, high values on **stringy**, low values on **striated**, moderately high values on **skewed**, medium values on **skinny**, and values across the full range on **splines**.


### iii.

Pairs of variables that have high values on skewed tend to have moderately low values on **outlying**, high values on **stringy**, high values on **striated**, and low values on **splines**.


### iv.

**marksInside50** and **disposal** is in outlier regarding the *clumpy* and *clumpy2* variables since it has high values for both variables, while for most other variable combinations, the values for *clumpy2* are very low or zero.
